### 2021-08-03 article/article.tex
update

To propagate the neutrino states through the Earth, we solve the Schrödinger equation with the Hamiltonian from Eq.~\ref{eq:H_NSI}. 
The Earth density profile is taken from the PREM~\cite{PREM}. The baseline for a given trajectory is determined using an average neutrino
production height of 15 km and an Earth radius of 6371 km. The propagation does not include neutrino absorption. After this, we are
ready to study the NSI effect on probability level.
%TODO: add more here? are there more effects than just absoption missing? justify absorption missing?

@@

\subsection{NSI phenomenology at flux level}\label{sec:nsiFluxEffects}
To study the NSI flux effect, we propagate the atmospheric neutrino flux from Honda et al~\cite{hondapaper} through the Earth.
The oscillation probability $P_{\alpha \beta}$ acts as a weight to the atmospheric flux, yielding the propagated flux for flavor $\beta$ at detector level as 
\begin{align}\label{eq:propFlux}
    \phi_\beta^\text{det} = \sum_\alpha P_{\alpha\beta} \phi_\alpha^\text{atm} \,,
\end{align}
where we sum over the initial lepton flavors $\alpha \in \{e,\mu, \bar{e}, \bar{\mu}\}$. To illustrate the impact of $\emt$ at flux level, 
we plot in Fig.~\ref{fig:flux_ratio} the quantity $(\phi_{\nu_\mu}^{NSI} + \phi_{\bar\nu_\mu}^{NSI})/(\phi_{\nu_\mu}^{SI} + \phi_{\bar\nu_\mu}^{SI})$, 
where all fluxes are propagated to detector level by Eq.~\ref{eq:propFlux}. 
In the left panel, we plot the flux in region in which 99\% of the 
DeepCore track events originating from $\nm + \anm$ fluxes are contained. In the right panel, we show the same but with IceCube events. We see that the only clear signal discernible to the IceCube detector
is a energy-independent flux deficiency with a factor in the order of $\sim 10^{0.5}$ from core-crossing neutrinos within a zenith range of $\ztrue < -0.87$. DeepCore on the other hand, 
is exposed to multiple fringes of flux surpluses with a factor in the order of $\sim 10$. The strongest surplus at \SI{20}{\GeV} is very weakly zenith dependent, a stark contrast to the
energy-independent but zenith-sensitive IceCube deficiency.
%TODO: check values above after final plot of flux ratios

@@

\begin{figure}[!tb]
   \begin{center}
      \includegraphics[width=0.8\linewidth]{figures/flux_ratio.pdf}
   \end{center}
   \caption{Ratio of detector NSI to SI atmospheric $\nu_\mu + \bar{\nu}_\mu$ fluxes expressed as $(\phi_{\nu_\mu}^{NSI} + \phi_{\bar\nu_\mu}^{NSI})/(\phi_{\nu_\mu}^{SI} + \phi_{\bar\nu_\mu}^{SI})$.
   We set $\emt = -0.05$, and all other $\epsilon_{\alpha\beta}=0$. 
   Left (right) panel shows the flux ratio in the energy range in which 99\% of the DeepCore (IceCube) events are contained.
   }\label{fig:flux_ratio}
\end{figure}%TODO: make prettier cmap

\section{Detector formalism}
\subsection{IceCube}\label{ch:ICmethod}
We obtain the data from the IC-86 sterile data release~\cite{IC2020}, which contain muon track events collected over 8 years. 
The reconstructed energy $\Ereco$ is logarithmically binned between \SI{500}{\GeV} to \SI{9976}{\GeV}, totalling 13 bins (index $i$).
The reconstructed cosine-zenith $\zreco$ is linearly binned between $-1$ and $0$ in 20 bins (index $j$).

The event count for each bin reads
\begin{align}\label{eq:ICevents}
   N_{ij} &= T\sum_\beta \int_{(\zreco)_j}^{(\zreco)_{j+1}} \dd \zreco \int_{\Ereco_{i}}^{\Ereco_{i+1}} \dd \Ereco \int_0^\pi R(\theta^R,\theta^T) \dd \ztrue \int_0^\infty \phi_\beta^\text{det}(\Etrue,\theta^R)  A^\text{eff}(\Etrue) R(\Ereco,\Etrue) 
   \dd \Etrue\,,
\end{align}
where $T$ is the live time of the detector, and $A^\text{eff}$ its effective area obtained from~\cite{ICaeff}, in which $A^\text{eff}$ has been averaged over neutrinos and 
anti-neutrinos. $R(x^R,x^T)$ is a resolution function, 
which is responsible for the smearing between the reconstructed and true parameters $x^R$ and $x^T$, respectively. We assume a log-normal distribution, giving it the form 
\begin{align}
    R(x^R, x^T) = \frac{1}{\sqrt{2\pi} \sigma_{x^R}x^R} \exp\left[-\frac{(\log x^R-\mu(x^T))^2}{2\sigma_{x^R}^2}\right]\,.
\end{align}
However, the energy reconstruction is biased, which means that we don't assume that $\mu(\Etrue) =\Ereco$~\cite{weaverThesis}. 
To model this relationship between $\Etrue$ and $\Ereco$, we train a Gaussian process regressor on the IC-86 Monte Carlo dataset from~\cite{IC2016}, from which
we can extract a predicted mean and standard deviation for a given $\Ereco$. We then take the $\Etrue$ points of the 99th percentile of each distribution to obtain
the limits over which to integrate. We take the angular resolution function to be identically unity since the angle resolution in IceCube for track-like 
events is less than $\SI{2}{\degree}$, making $\ztrue$ practically coincide with $\zreco$ for our study~\cite{IC2020}.

@@

The data include 14901 track-like events and 26001 cascade-like events, both divided into eight 
$ \log_{10}E^{reco} \in [0.75,1.75]$ bins (index $i$), eight $\zreco \in [-1,1]$ bins (index $j$), and two bins for the event classification (index $k$). 
Each event has a Monte Carlo weight $w_{ijk,\beta}$ from which we can construct the event count as
\begin{align}\label{eq:MCevents}
    N_{ijk} &= C_{ijk}\sum_{\beta}w_{ijk,\beta}\, \phi_\beta^\text{det}\,,
\end{align}
where $C_{ijk}$ is the correction factor from the detector systematic uncertainty and $\phi_\beta^\text{det}$ is defined as Eq.~\ref{eq:propFlux}. We have now substituted the effect of the Gaussian smearing 
by treating the reconstructed and true quantities as a migration matrix.

\subsection{PINGU}\label{ch:PINGUmethod}
The methodology behind the PINGU simulations is the same as with our DeepCore study~\ref{ch:DCmethod}. We use the public Monte Carlo~\cite{PINGUdata}, 
which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modeled for PINGU, the correction factors $C_{ijk}$ are all unity. We will remedy this by including an uncorrelated systematic error.
As with the DeepCore Monte Carlo, the PINGU Monte Carlo is divided into eight 
$\log_{10}E^{reco} \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We plot the normalized event differences $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ for cascades and tracks in Fig.~\ref{fig:event_pulls}. %TODO:talk about this
We generate `data' by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

@@

\section{Results}
At the reconstructed event level, we note that the $\emt$ features discussed in Sec.~\ref{sec:nsiEffects} display themselves differently in each detector.
We investigate this for $\emt$ by in Fig.~\ref{fig:Pmm_asymmetry} plotting the muon survival probability difference $P_{\mu\mu}(\epsilon^+_{\mu\tau}) - P_{\mu\mu}(\epsilon^-_{\mu\tau})$, where 
%TODO: check that this is what is plotted.
$\epsilon^\pm_{\mu\tau} = \pm 0.01$.
Comparing Fig.~\ref{fig:Pmm_asymmetry} with the flux ratios in Fig.~\ref{fig:emt_events}, we see that the reconstruction of PINGU is
superior to DeepCore, since the event ratio $\log{(N^+_{NSI}/N^-_{NSI})} = \log{N^+_{NSI}} - \log{N^-_{NSI}}$ (in reconstructed quantities) 
more closely matches the probability difference $P^+-P^-$ (in true quantities).
This is most evident below \SI{20}{\GeV}, where the DeepCore reconstruction has washed out the fringes while PINGU preserves the $N^-_{NSI}$ surplus below \SI{10}{\GeV}.

@@

\subsection{Constraining the NSI parameters}\label{sec:method}
In this section, we will constrain the four NSI parameters $\ett$, $\emt$, $\eem$, and $\eet$ by considering the detectors separately as well as jointly.
For our analyses, we define our $\chi^2$ as
\begin{align} \label{eq:chisq}
    \chi^{2}(\hat{\theta},\alpha,\beta, \kappa)=\sum_{ijk} \frac{\left(N^\text{th}-N^\text{data}\right)_{ijk}^{2}}
    {\left(\sigma^\text{data}_{ijk}\right)^{2} + \left(\sigma^\text{syst}_{ijk}\right)^{2}}+ 
    \frac{(1-\alpha)^2}{\sigma_\alpha^2} + \frac{\beta^2}{\sigma_\beta^2}\,
\end{align}
where we minimize over the model parameters $\hat{\theta} \in \{\dm, \theta_{23}, \epsilon\}$, the penalty terms $\alpha$ and $\beta$, and the free parameter $\kappa$.
$N_{ijk}^\text{th}$ is the expected number of events from theory, and $N_{ijk}^\text{data}$ is the observed number of events in that bin.

In our simulations of $N_{ijk}^\text{th}$, we set all standard oscillation parameters to their current best-fit values of Eq.~\ref{eq:PINGUparams}, 
except for $\dm$ and $\theta_{23}$
which we let free over their $3\sigma$ limits \SIrange{2.435e-3}{2.598e-3}{\eV \squared} and \SIrange{40.1}{51.7}{\degree}, respectively.

@@

We treat the uncorrelated systematic uncertanties differently for each detector. For IceCube, we set $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$.
We consider best, normal, and worst-case scenarios in IceCube using
$f=5\%$, $10\%$, and $15\%$ respectively. For PINGU, we use the same form but instead use $f=0\%$, $3\%$, and $5\%$. %TODO: maybe source/justification for these values?
For DeepCore, we use the provided systematic error distribution which accounts for uncertainties in the finite MC statistics and the data-driven 
muon background estimate~\cite{DC2019data}. This is summarized in Table~\ref{table:syst_errors}.  %TODO segue
{\renewcommand{\arraystretch}{1.2}
\begin{table}
   \begin{tabular}{lccc}
      \hline \hline
      Experiment & Best case & Baseline & Worst case \\
      \hline
      IceCube & $5\%$ & $10\%$ & $15\%$ \\
      PINGU & $0\%$ & $3\%$ & $5\%$ \\
      \hline \hline
   \end{tabular}
   \caption{Definition of the best, baseline, and worst case scenarios considered in each experiment, modelled by $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$ with $f$ from the table.
   We do not consider different DeepCore scenarios because her systematic error distribution is already provided in the data release~\cite{DC2019data}.}\label{table:syst_errors}
\end{table}

>>>>

To propagate the neutrino states through the Earth, we solve the Schrödinger equation with the Hamiltonian from Eq.~\ref{eq:H_NSI}. 
The Earth density profile is taken from the PREM~\cite{PREM}, and we do not consider neutrino absorption.
The baseline for a given trajectory is determined using an average neutrino production height of 15 km and an Earth radius of 6371 km.
Thus, matter oscillations introduce a zenith angle dependence in two areas. Firstly, the matter potential depends on the matter density through the
neutrino propagates. A neutrino with $\cos{\theta_z} < -0.87$ will traverse through the dense Earth core, whereas a mantle-crossing neutrino 
with $\cos{\theta_z} > -0.87$ will not. Secondly, the zenith angle also alters the baseline through which a neutrino undergoes matter oscillations.
After this, we are ready to study the NSI effect on probability level.
%TODO: add more here? are there more effects than just absoption missing? justify absorption missing?

@@

\subsection{NSI phenomenology at flux level}\label{sec:nsiFluxEffects}
To study the NSI flux effect, we propagate the atmospheric neutrino flux from Honda et al~\cite{hondapaper} through the Earth.
The oscillation probability $P_{\alpha \beta}$ acts as a weight to the atmospheric flux, yielding the propagated flux for flavor $\beta$ at detector level as 
\begin{align}\label{eq:propFlux}
    \phi_\beta^\text{det} = \sum_\alpha P_{\alpha\beta} \phi_\alpha^\text{atm} \,,
\end{align}
where we sum over the initial lepton flavors $\alpha \in \{e,\mu, \bar{e}, \bar{\mu}\}$.
To illustrate the impact of $\emt$ at flux level, 
we plot in Fig.~\ref{fig:flux_ratio} the quantity $(\phi_{\nu_\mu}^{NSI} + \phi_{\bar\nu_\mu}^{NSI})/(\phi_{\nu_\mu}^{SI} + \phi_{\bar\nu_\mu}^{SI})$, 
where all fluxes are propagated to detector level by Eq.~\ref{eq:propFlux}. 
In the left (right) panel, we plot the flux in region in which 99\% of the 
DeepCore (IceCube) track events originating from $\nm + \anm$ fluxes are contained. We see that the only clear signal discernible to the IceCube detector
is a energy-independent flux deficiency with a factor in the order of $\sim 10^{0.5}$ from core-crossing neutrinos within a zenith range of $\ztrue < -0.87$. DeepCore on the other hand, 
is exposed to multiple fringes of flux surpluses with a factor in the order of $\sim 10$. The strongest surplus at \SI{20}{\GeV} is very weakly zenith dependent, a stark contrast to the
energy-independent but zenith-sensitive IceCube deficiency.
%TODO: check values above after final plot of flux ratios

@@

\begin{figure}[!tb]
   \begin{center}
      \includegraphics[width=0.8\linewidth]{figures/flux_ratio.pdf}
   \end{center}
   \caption{Ratio of detector NSI to SI atmospheric $\nu_\mu + \bar{\nu}_\mu$ fluxes expressed as $(\phi_{\nu_\mu}^{NSI} + \phi_{\bar\nu_\mu}^{NSI})/(\phi_{\nu_\mu}^{SI} + \phi_{\bar\nu_\mu}^{SI})$.
   We set $\emt = -0.05$, and all other $\epsilon_{\alpha\beta}=0$. 
   Left (right) panel shows the flux ratio in the energy range in which 99\% of the simulated DeepCore (IceCube) events are contained.
   }\label{fig:flux_ratio}
\end{figure}%TODO: make prettier cmap

\section{Detector formalism}
The IceCube Neutrino Observatory is a gigaton Cherenkov detector located within Antarctic ice near the Geographic 
South Pole. It consists of 5160 individual digital optical modules situated on 86 strings, which detect Cherenkov light emitted by charged leptons
from neutrino events. We distinguish between two interaction topologies: track- and cascade-like events. Track-like events originate mainly
from muons originating from $\nm$ CC interactions, while cascades predominantly consist of electromagnetic and hadronic showers.

DeepCore is a dense sub-array whithin the standard IceCube array, which allows us to observe neutrino interactions down to a few GeV~\cite{DC2021}.
Due to its lower energy threshold compared with IceCube, DeepCore has previously been used to study NSI effects in this more discernible range\cite{DC2021,deepcoreNSI}.

PINGU is a proposed 26 string in-fill array to IceCube, consisting of optical modules of similar nature~\cite{PINGUletter}. While porposed to mainly shed light on 
neutrino mass ordering, PINGU will be able to probe NSI parameters aswell.  
\subsection{IceCube}\label{ch:ICmethod}

We obtain the data from the IC-86 sterile data release~\cite{IC2020}, which contain muon track events collected over 8 years. 
In the data, the reconstructed energy $\Ereco$ is logarithmically binned between \SI{500}{\GeV} to \SI{9976}{\GeV}, totalling 13 bins (index $i$).
The reconstructed cosine-zenith $\zreco$ is linearly binned between $-1$ and $0$ in 20 bins (index $j$).

The event count for each bin reads
\begin{align}\label{eq:ICevents}
   N_{ij} &= T\sum_\beta \int_{(\zreco)_j}^{(\zreco)_{j+1}} \dd \zreco \int_{\Ereco_{i}}^{\Ereco_{i+1}} \dd \Ereco \int_0^\pi R(\theta^R,\theta^T) \dd \ztrue \int_0^\infty \phi_\beta^\text{det}(\Etrue,\theta^R)  A^\text{eff}(\Etrue) R(\Ereco,\Etrue) 
   \dd \Etrue\,,
\end{align}
where $T$ is the live time of the detector, and $A^\text{eff}$ its effective area obtained from~\cite{ICaeff}, in which $A^\text{eff}$ has been averaged over neutrinos and 
anti-neutrinos. $R(x^R,x^T)$ is a resolution function, 
which is responsible for the smearing between the reconstructed and true parameters $x^R$ and $x^T$, respectively. We assume a log-normal distribution, giving it the form 
\begin{align}
    R(x^R, x^T) = \frac{1}{\sqrt{2\pi} \sigma_{x^R}x^R} \exp\left[-\frac{(\log x^R-\mu(x^T))^2}{2\sigma_{x^R}^2}\right]\,.
\end{align}
However, the energy reconstruction is biased, meaning that the most probable reconstructed energy for a given true energy is not the same due to 
systematic differences~\cite{weaverThesis}. Thus, we don't assume that $\mu(\Etrue) =\Ereco$. 
To model this relationship between $\Etrue$ and $\Ereco$, we fit a Gaussian process regressor on the IC-86 Monte Carlo dataset from~\cite{IC2016}, from which
we can extract a predicted mean and standard deviation for a given $\Ereco$. We then take the $\Etrue$ points of the 99th percentile of each reconstructed
distribution to obtain the $\Etrue$ limits over which to integrate. We take the angular resolution function to be identically unity since the angle resolution in IceCube for track-like 
events is less than $\SI{2}{\degree}$, making $\ztrue$ practically coincide with $\zreco$ for our study~\cite{IC2020}.

@@

The data include 14901 track-like events and 26001 cascade-like events, both divided into eight 
$ \log_{10}\Ereco \in [0.75,1.75]$ bins (index $i$), eight $\zreco \in [-1,1]$ bins (index $j$), and two bins for the event classification (index $k$). 
Each event has a Monte Carlo weight $w_{ijk,\beta}$ from which we can construct the event count as
\begin{align}\label{eq:MCevents}
    N_{ijk} &= C_{ijk}\sum_{\beta}w_{ijk,\beta}\, \phi_\beta^\text{det}\,,
\end{align}
where $C_{ijk}$ is the correction factor from the detector systematic uncertainty and $\phi_\beta^\text{det}$ is defined as Eq.~\ref{eq:propFlux}. 
We have now substituted the effect of the Gaussian smearing by treating the reconstructed and true quantities as a migration matrix.

\subsection{PINGU}\label{ch:PINGUmethod}
The methodology behind the PINGU simulations is the same as with our DeepCore study~\ref{ch:DCmethod}. We use the public Monte Carlo~\cite{PINGUdata}, 
which allows us to construct the event count as in Eq.~\ref{eq:MCevents}.
However, since no detector systematics is yet modeled for PINGU, we take the correction factors $C_{ijk}$ are all unity. We will remedy this by including an uncorrelated systematic error,
which we can scale at will.
As with the DeepCore Monte Carlo, the PINGU Monte Carlo is divided into eight 
$\log_{10}\Ereco \in [0.75,1.75]$ bins, and eight $\zreco \in [-1,1]$ bins for both track- and cascade-like events. 
We plot the normalized event differences $(N_{NSI} - N_{SI})/\sqrt{N_{SI}}$ for cascades and tracks in Fig.~\ref{fig:event_pulls}. %TODO:talk about this
We generate `data' by predicting the event rates at PINGU with the following best-fit parameters from~\cite{nufit}, except for the CP-violating phase which is set to zero for simplicity.

@@

\section{Results}
At the reconstructed event level, we note that the $\emt$ features discussed in Sec.~\ref{sec:nsiFluxEffects} display themselves differently in each detector.
We investigate this for $\emt$ by in Fig.~\ref{fig:Pmm_asymmetry} plotting the muon survival probability difference $P_{\mu\mu}(\epsilon^+_{\mu\tau}) - P_{\mu\mu}(\epsilon^-_{\mu\tau})$, where 
%TODO: check that this is what is plotted.
$\epsilon^\pm_{\mu\tau} = \pm 0.01$.
Comparing Fig.~\ref{fig:Pmm_asymmetry} with the flux ratios in Fig.~\ref{fig:emt_events}, we see that the reconstruction of PINGU is
superior to DeepCore, since the event ratio $\log{(N^+_{NSI}/N^-_{NSI})} = \log{N^+_{NSI}} - \log{N^-_{NSI}}$ (in reconstructed quantities) 
more closely matches the probability difference $P^+-P^-$ (in true quantities).
This is most evident below \SI{20}{\GeV}, where the DeepCore reconstruction has washed out the fringes while PINGU preserves the $N^-_{NSI}$ surplus below \SI{10}{\GeV}.

@@

\subsection{Constraining the NSI parameters}\label{sec:constraining}
In this section, we will constrain the four NSI parameters $\ett$, $\emt$, $\eem$, and $\eet$ by considering the detectors separately as well as jointly.
For our analyses, we define our $\chi^2$ as
\begin{align} \label{eq:chisq}
    \chi^{2}(\hat{\theta},\alpha,\beta, \kappa)=\sum_{ijk} \frac{\left(N^\text{th}-N^\text{data}\right)_{ijk}^{2}}
    {\left(\sigma^\text{data}_{ijk}\right)^{2} + \left(\sigma^\text{syst}_{ijk}\right)^{2}}+ 
    \frac{(1-\alpha)^2}{\sigma_\alpha^2} + \frac{\beta^2}{\sigma_\beta^2}\,
\end{align}
where we minimize over the model parameters $\hat{\theta} \in \{\dm, \theta_{23}, \epsilon\}$, the penalty terms $\alpha$ and $\beta$, and the free parameter $\kappa$.
$N_{ijk}^\text{th}$ is the expected number of events from theory, and $N_{ijk}^\text{data}$ is the observed number of events in that bin.

In our simulations of $N_{ijk}^\text{th}$, we set all standard oscillation parameters to their current best-fit values of Eq.~\ref{eq:PINGUparams}, 
except for $\dm$ and $\theta_{23}$
which we vary over their $3\sigma$ limits \SIrange{2.435e-3}{2.598e-3}{\eV \squared} and \SIrange{40.1}{51.7}{\degree}, respectively.

@@

We treat the uncorrelated systematic uncertanties differently for each detector. For IceCube, we set $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$.
We consider best, normal, and worst-case scenarios in IceCube using
$f=5\%$, $10\%$, and $15\%$ respectively. For PINGU, we use the same form but instead use $f=0\%$, $3\%$, and $5\%$. %TODO: maybe source/justification for these values?
For DeepCore, we use the provided systematic error distribution which accounts for uncertainties in the finite MC statistics and the data-driven 
muon background estimate~\cite{DC2019data}. This is summarized in Table~\ref{table:syst_errors}.  %TODO segue
{\renewcommand{\arraystretch}{1.2}
\begin{table}
   \begin{tabular}{lccc}
      \hline \hline
      Experiment & Best case & Baseline & Worst case \\
      \hline
      IceCube & $5\%$ & $10\%$ & $15\%$ \\
      PINGU & $0\%$ & $3\%$ & $5\%$ \\
      \hline \hline
   \end{tabular}
   \caption{Our definition of the best, baseline, and worst case scenarios considered in each experiment, modelled by $\sigma_{ijk}^\text{syst} = f\sqrt{N_{ijk}^\text{data}}$ with $f$ from the table.
   We do not consider different DeepCore scenarios because her systematic error distribution is already provided in the data release~\cite{DC2019data}.}\label{table:syst_errors}
\end{table}
